---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# greta.car

<!-- badges: start -->
<!-- once you've signed into travis and set it to wath your new repository, you can edit the following badges to point to your repo -->
[![Codecov test coverage](https://codecov.io/gh/greta-dev/greta.car/branch/main/graph/badge.svg)](https://codecov.io/gh/greta-dev/greta.car?branch=main)
[![R-CMD-check](https://github.com/goldingn/greta.car/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/goldingn/greta.car/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This is an experimental extension package to greta for implementing conditional autoregressive models. Currently implemented is an experimental `icar()` distribution for efficiently modelling intrisic conditional autoregressive models, and an example analysis for it

```{r echo = FALSE}
set.seed(2023-05-19)
```

Define an ICAR model on a network
```{r}
# define a network on a ring with a given number of nodes
n_nodes <- 20
edges <- cbind(1:n_nodes, c(2:n_nodes, 1))
```

Simulate some true parameters and fake Gaussian observations at each
```{r}
true_phi <- rnorm(n_nodes)
true_phi <- true_phi - sum(true_phi) / n_nodes
# sum(true_phi)
true_intercept <- 2
true_sd <- 0.2
y <- rnorm(n = n_nodes,
           mean = true_intercept + true_phi,
           sd = true_sd)
```

Visualise this data with igraph
```{r message = FALSE}
library(igraph)
g <- graph_from_edgelist(edges, directed = FALSE)
scale <- y
scale <- scale - min(scale)
scale <- scale / max(scale)
V(g)$color <- grey(scale)
plot(g)
```


Define a greta model for this
```{r}
library(greta.car)

# define an ICAR random effect on this network
phi <- icar(edges)
# phi has length n_nodes (extracted from 'edges')

# apply soft sum-to-zero, to avoid identifiability issues
nil <- zeros(1)
distribution(nil) <- normal(sum(phi), n_nodes * 0.001)

# define an observation model for the data
intercept <- normal(0, 1)
sd <- normal(0, 1, truncation = c(0, Inf))
distribution(y) <- normal(intercept + t(phi), sd)
```
Fit the model by MCMC
```{r}
m <- model(phi)
draws <- mcmc(m)
```
Check convergence of chains (results hidden here)
```{r results = "hide", fig.show = "hide"}
coda::gelman.diag(draws, autoburnin = FALSE, multivariate = FALSE)
plot(draws)
```

Plot concordance of phi estimates with 'true' values
```{r}
estimate <- summary(draws)$statistics[, "Mean"]
ci <- summary(draws)$quantiles[, c("2.5%", "97.5%")]
truth <- c(true_phi)
plot(estimate ~ truth,
     ylim = range(ci))
arrows(x0 = truth,
       x1 = truth,
       y0 = ci[, 1],
       y1 = ci[, 2], length = 0)
abline(0, 1, lty = 2)
```

Visualise the posterior means with iGraph (note the orders and locations of nodes have changed from the above)
```{r}
library(igraph)
g <- graph_from_edgelist(edges, directed = FALSE)
scale <- estimate
scale <- scale - min(scale)
scale <- scale / max(scale)
V(g)$color <- grey(scale)
plot(g)
```

---
 
The package infrastructure itself is from the [greta.template](https://github.com/greta-dev/greta.template) repo, and checklist for setting that up remains below.

#### to do list

 - [x] Pick a package name (preferably with 'greta.' at the beginning). Update the package name in:
    - [x] the 'Package' field in  `DESCRIPTION`
    - [x] the `library()` and `test_check()` calls in `tests/testthat.R`
    - [x] in the `@name` documentation field in `R/package.R`
    - [x] at the top of `README.Rmd`
    - [x] the repo name (if it changed since you made it)
    - [x] the R project file; rename it from `greta.template.Rproj` to `greta.<something>.Rproj`
 - [x] Come up with a helpful package title. Add it to:
    - [x] the 'Title' field in `DESCRIPTION`
    - [x] the `@title` documentation field in `R/package.R`
    - [x] at the top of your GitHub repo 
 - [x] Fill in the the author details in `DESCRIPTION`
 - [x] Update the 'URL' and 'BugReports' fields in `DESCRIPTION` to point to your repo
 - [x] Decide what sort of license you want to use for your package (you are completely free to change the CC0 license in the template). See [https://choosealicense.com]() or  [r-pkgs.org](https://r-pkgs.org/description.html#license) for help choosing.
    - [x] create the license file and edit the 'License' field in `DESCRIPTION` (e.g. with [the `usethis` package](https://usethis.r-lib.org/reference/licenses.html))
 - [x] Write short paragraph describing the package. Copy it to:
    - [x] the 'Description' field of `DESCRIPTION`
    - [x] the `@description` documentation field of `R/package.R` 
 - [x] add a github actions badge with `use_github_actions_badge()`
 - [x] edit the [codecov](https://codecov.io) badge in `README.Rmd` to point to your package
 - [x] Write a simple example introducing the package and add it to `R/package.R`
 - [x] Update `README.Rmd` to sell to people with a sales pitch and maybe an example that creates a figure, to get people excited about your package.
 - [x] Start adding functions, documentation and examples to new R files in `R` folder
 - [ ] Write some unit tests for these functions in the `tests/testthat` folder
 - [x] Delete the example function and test files `R/square.R` and `tests/testthat/test-square.R` 
 
